\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2017-06-02 eg6-letters/NIM/EG6_RTPC_NIM.tex
some cleaning

\begin{figure}[tbp]
\centering
\includegraphics[scale=0.70]{fig/GEM_photo.jpg}
\caption{Picture of a GEM foil similar to the one used for our RTPC. {\color{red} Source?}} 
\label{fig:GEMs}
\end{figure}

The GEM technology has been chosenfor its flexibility as it
allows for a curved amplification surface. Also, GEMs are known to rarely create 
sparks~\cite{GEM_ref}, which is important when trying to detect 
highly ionizing slow nuclei that deposit large amount of energy. The GEMs for 
this RTPC are made from a Kapton insulator layer, 50~$\mu$m 
thick, sandwiched between two 5~$\mu$m copper layers. The mesh of each GEM 
layer is chemically etched with 50~$\mu$m diameter holes with double-conical 
shapes as illustrated in figure~\ref{fig:GEMs}. The potential difference 
applied between the two copper layers of the GEM creates a very strong 
electric field in each hole leading to high ionization and amplification.

The drift gas used in the experiment is a 80-20\% neon-DME mixture, this choice 
has been made in order to balance the energy deposit with a reasonnable
Lorentz angle. Calculations using the MAGBOLTZ program~\cite{MAGBOLTZ} 
showed that with the perpendicular 5~T field, we would have a Lorentz angle of 
about 23$^\circ$ with this gas mixture. This is already a significant angle,
a heavier gas would have larger angle increasing significantly
the length of the drift paths as the electron would 
make a long spiral around the beam line before to reach the GEMs.

One of the important problematic to develop the radial TPC was to obtain a 
good support structure of the GEM foils, allowing an easy installation of the 
GEM in good conditions and with high precision. At the same time, we wanted to 
keep the material budget small in the forward region where we detect other 
particles in subsequent detectors. We successfuly realized these prerequisites 
by using fiber glass rings glued to each ends of the GEM foils to form self 
supporting cylinders that could be installed independently in the RTPC after 
all gluing and soldering operations. The rigidity of the GEM foil is enough
for the structure to be selfsupporting and only the 
backward part of the cylinder was fixed to the main mechanical structure. This 
design only left a light fiberglass ring in the forward region, reducing to a 
minimum secondary interactions.

@@

In order to read all the detector readout pads, four FEC crates are used, 
each equipped with 6 boards, plus a ROC. A schematic of the readout system, 
for a single crate, is reported in figure~\ref{fig:ReadoutScheme}. This 
configuration permits to reduce the dead time associated to FEC readout 
operations, scaling linearly with the number of boards in the crate. During 
the 2009 run, the system was succesfully operated with a DAQ rate of 3.1 kHz, 
with a live time of $70 \%$.

@@

In this section we will detail the methods used to calibrate the drift speed,
drift paths and gains of the detector. Drift speed and paths were initially
calculated using the MAGBOLTZ~\cite{MAGBOLTZ} program, then refined using
data to account for variations of the run conditions. We always assume 
cylindrical symmetry in the chamber for the calibration, such that none of
the parameters depend on the asimuthal angle $\phi$. The initial MAGBOLTZ
calibration was improved through several iterations of the
process described below, each iteration increasing the number of events 
properly reconstructed in the RTPC. The figures presented in this section
are the one obtained while performing the last iteration of this long 
calibration process.

@@

The drift paths are ajusted using a set of identified elastic events 
from our lower beam energy run. Based on the kinematic of the electrons in
these events, we generate the helium nucleus in a GEANT4 
simulation~\cite{GEANT4} of our RTPC. Then we compare the calculated GEANT4 trajectory of 
the Helium nuclei to the hits measured in the chamber.

@@

The second is a coherent noise affecting about 25\% of the pramplifiers, where the signature is simultaneous hits in most of the 16 channels in a preamp group.   An event-based technique to identify and remove it was developed based on counting simultaneous hits in preamp group, and, if sufficiently large, perform a dynamic pedestal subtraction based on the average ADC of the channels' neighbors within the preamp group.

@@

We reported on the construction, operation and calibration of a small RTPC 
designed to measure helium-4 nuclei in high rate environnement. The operation
of the detector was successful and allowed to detect helium nuclei at a rate
of several kHz in coincidence with the detection of high energy electrons and 
photons in the CLAS spectrometer.

>>>>

\begin{figure}[tbp]
\centering
\includegraphics[scale=0.70]{fig/GEM_photo.jpg}
\caption{Picture of a typical GEM foil similar to the one used for our RTPC.  
{\color{red} Source?}} \label{fig:GEMs}
\end{figure}

The GEM technology has been chosen for its flexibility as it
allows for a curved amplification surface. Also, GEMs are known to rarely create 
sparks~\cite{GEM_ref}, which is important when trying to detect 
highly ionizing slow nuclei that deposit large amount of energy. The GEMs for 
this RTPC are made from a Kapton insulator layer, 50~$\mu$m 
thick, sandwiched between two 5~$\mu$m copper layers. The mesh of each GEM 
layer is chemically etched with 50~$\mu$m diameter holes with double-conical 
shapes as illustrated in figure~\ref{fig:GEMs}. The potential difference 
applied between the two copper layers of the GEM creates a very strong 
electric field in each hole leading to high ionization and amplification.

The drift gas used in the experiment is a 80-20\% neon-DME mixture, this choice 
has been made in order to balance the energy deposit with a reasonable
Lorentz angle. Calculations using the MAGBOLTZ program~\cite{MAGBOLTZ} 
showed that with the perpendicular 5~T field, we would have a Lorentz angle of 
about 23$^\circ$ with this gas mixture. This is already a significant angle,
a heavier gas would have larger angle increasing significantly
the length of the drift paths as the electron would 
make a long spiral around the beam line before to reach the GEMs.

One of the important problematic to develop the radial TPC was to obtain a 
good support structure of the GEM foils, allowing an easy installation of the 
GEM in good conditions and with high precision. At the same time, we wanted to 
keep the material budget small in the forward region where we detect other 
particles in subsequent detectors. We successfully realized these prerequisites 
by using fiber glass rings glued to each ends of the GEM foils to form self 
supporting cylinders that could be installed independently in the RTPC after 
all gluing and soldering operations. The rigidity of the GEM foil is enough
for the structure to be self-supporting and only the backward part of the 
cylinder was fixed to the main mechanical structure. This design only left a 
light fiberglass ring in the forward region, reducing to a minimum secondary 
interactions.

@@

In order to read all the detector readout pads, four FEC crates are used, 
each equipped with 6 boards, plus a ROC. A schematic of the readout system, 
for a single crate, is reported in figure~\ref{fig:ReadoutScheme}. This 
configuration permits to reduce the dead time associated to FEC readout 
operations, scaling linearly with the number of boards in the crate. During the 
2009 run, the system was successfully operated with a DAQ rate of 3.1 kHz, with 
a live time of $70 \%$.

@@

In this section we will detail the methods used to calibrate the drift speed,
drift paths and gains of the detector. Drift speed and paths were initially
calculated using the MAGBOLTZ~\cite{MAGBOLTZ} program, then refined using
data to account for variations of the run conditions. We always assume 
cylindrical symmetry in the chamber for the calibration, such that none of
the parameters depend on the azimuthal angle $\phi$. The initial MAGBOLTZ
calibration was improved through several iterations of the
process described below, each iteration increasing the number of events 
properly reconstructed in the RTPC. The figures presented in this section
are the one obtained while performing the last iteration of this long 
calibration process.

@@

The drift paths are adjusted using a set of identified elastic events from our 
lower beam energy run. Based on the kinematic of the electrons in these events, 
we generate the helium nucleus in a GEANT4 simulation~\cite{GEANT4} of our 
RTPC. Then we compare the calculated GEANT4 trajectory of the Helium nuclei to 
the hits measured in the chamber.

@@

The second is a coherent noise affecting about 25\% of the preamplifiers, where 
the signature is simultaneous hits in most of the 16 channels in a preamp 
group.   An event-based technique to identify and remove it was developed based 
on counting simultaneous hits in preamp group, and, if sufficiently large, 
perform a dynamic pedestal subtraction based on the average ADC of the 
channels' neighbors within the preamp group.

@@

We reported on the construction, operation and calibration of a small RTPC 
designed to measure helium-4 nuclei in high rate environment. The operation
of the detector was successful and allowed to detect helium nuclei at a rate
of several kHz in coincidence with the detection of high energy electrons and 
photons in the CLAS spectrometer.
